---
title: | 
  | \fontsize{24}{24}\sf \color{yellow} Missing Data
  | \fontsize{20}{20}\sf Data Quality:
author: "Alfonso R. Reyes"
date: "April 2019"
fontsize: 14pt
output:
  beamer_presentation: 
    fig_height: 3.5
    fig_width: 7
    highlight: kate
    includes:
      in_header: headerOG.tex
    theme: AnnArbor
---


```{r setup-MD, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      comment = "#>",
                      cache = TRUE
                      )
options(digits=4)
```

# Finding and filling missing data

## Example of filling missing data

\begin{textblock}{4.5}(4.5, 4.5)\textcolor{black}{\fontsize{28}{28}{\textbf{Well Tests}}}\end{textblock} 


```{r echo=FALSE}
# tibble allows to compress the output of a dataset when printing
library(dplyr)
library(tibble)
library(skimr)

dataset <- read.csv("./inst/rawdata/2007_welltest.csv", stringsAsFactors = FALSE)
dataset <- as_tibble(dataset)   # small output
```


```{r echo=FALSE}
# show all the columns 
# names(dataset)
```

## Well test data: numeric variables
\fontsize{7}{7}\sf
```{r}
# get only the numeric columns
nums <- sapply(dataset, is.numeric)
numdf <- dataset[, nums]
names(numdf)
```

```{r echo=FALSE}
# numdf
```


```{r echo=FALSE}
sel_numdf <- 
numdf %>% 
  select(DURATION_HRS, WH_USC_PRESS_PSIG, ANNULUS_PRESS_PSIG, WH_USC_TEMP_F, 
         GL_RATE_SM3PERDAY, NET_OIL_RATE_ADJ_SM3PERDAY, GOR_SCFPERBBL, 
         WATERCUT_PCT)
    # as.data.frame(sapply(as.numeric)) %>% 
```

```{r echo=FALSE}
# get only the non-numeric columns
nums <- sapply(dataset, is.numeric)
nonumdf <- dataset[, !nums]
names(nonumdf)
```



```{r echo=FALSE}
# show only the ID descriptive variables
library(dplyr)

id_descr <- 
dataset %>%
    select(CODE, WELL_HOLE_CODE, OP_SUB_AREA_CODE, OP_AREA_CODE, OP_FCTY_1_CODE)
```

```{r echo=FALSE}
# bind the columns from the non-numeric and numeric dataframes
df <- cbind(id_descr$CODE, id_descr$WELL_HOLE_CODE, sel_numdf)
```


## Look for missing values

\begin{alertblock}{}
In this case we are using the statistical package \textbf{mice} to fill in for the missing values.
\end{alertblock}  

\vspace{1.2cm}


\fontsize{10}{10}\sf

```{r echo=TRUE}
library(tidyverse)
library(mice)
library(lattice)
```

```{r echo=TRUE, eval=FALSE}
# plot the missing data in a matrix by variables
md_pattern <- md.pattern(df, rotate.names = TRUE)
```


## Look for missing values


\fontsize{10}{10}\sf

```{r fig.height=13, fig.width=15}
# plot the missing data in a matrix by variables
md_pattern <- md.pattern(df, rotate.names = TRUE)
```


## Missing values per variable
\fontsize{7}{7}\sf

```{r echo=TRUE}
# number of complete cases: no variable value is missing in that observation
sum(complete.cases(sel_numdf))
```

\vspace{1.2cm}

```{r echo=TRUE}
sel_numdf %>% 
    skim_to_list() %>% 
    .[["numeric"]] %>% 
    as.data.frame(sapply(as.numeric)) %>% 
    dplyr::select(variable, missing, complete)
```


```{r imputting_data, echo=FALSE, cache=TRUE}
# create mice object dataset_impute
df_impute <- mice(data = df[, -c(1,2)],  print = FALSE, m = 1)
```


## How well the fill data fits in

```{r}
# magenta is the imputted values
stripplot(df_impute, pch = 20, cex = 1.2, 
          par.strip.text=list(cex=0.55))
```




